---
title: "Transmission effect on car MPG performance"
author: "Henrique Souza"
date: "14/02/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Executive Summary

Motor Trend Magazine is interested in exploring the relationship between a set of variables and miles per gallon (MPG). They are particularly interested in the following two questions:

1. Is an automatic or manual transmission better for MPG.
1. Quantify the MPG difference between automatic and manual transmissions.

In this document we compare different models and show that *for an average car, manual transmission outperforms automatic in <strong>11%</strong> MPG efficience*.

## Exploratory Data Analysis

You can see a comparison between Automatic and Manual cars accross diferent dimensions
using a small Shiny Application available on the following url:

https://htssouza.shinyapps.io/DevelopingDataProducts-Project/

The source of this application and of this presentation itself can be found on:

https://github.com/htssouza/DevelopingDataProducts-Project/

## Models

The following 5 linear models were compared:

1. Using only the Transmission.
1. Transmission and Weight.
1. Transmission, Weight and HP.
1. Transmission, Weight, HP and Displacement.
1. Transmission, Weight, HP, Displacement and Gears.

```{r echo = FALSE, message = FALSE}
library(datasets)
library(dplyr)
library(ggplot2)
library(stats)
data(mtcars)
am.levels <- c("auto", "manual")
am.auto <- as.factor(c(1, 2))
am.manual <- as.factor(c(1, 2))
levels(am.auto) <- am.levels
levels(am.manual) <- am.levels
am.auto <- am.auto[1]
am.manual <- am.manual[2]
mtcars <- mtcars %>%
  mutate(am = am + 1) %>%
  mutate(am = factor(am))
levels(mtcars$am) <- am.levels

mod1 <- lm(mpg ~ am, data = mtcars)
mod2 <- lm(mpg ~ am + wt, data = mtcars)
mod3 <- lm(mpg ~ am + wt + hp, data = mtcars)
mod4 <- lm(mpg ~ am + wt + hp + disp, data = mtcars)
mod5 <- lm(mpg ~ am + wt + hp + disp + gear, data = mtcars)
# anova(mod1, mod2, mod3, mod4, mod5)
```

The p-values for Model 4 is already too big (>= 0.05).
We will stick with Model 3 (Transmission, Weight and HP).

## Conclusion

Using our model to compare the performance of average cars with both transmission
types we have that a *Manual* one has a better MPG performance of:

```{r echo = TRUE}
wt.mean <- mean(mtcars$wt)
hp.mean <- mean(mtcars$hp)
auto.mpg <- predict(mod3, data.frame(am = am.auto, wt = wt.mean, hp = hp.mean))
manual.mpg <- predict(mod3, data.frame(am = am.manual, wt = wt.mean, hp = hp.mean))
cat(paste(round(as.numeric(((manual.mpg - auto.mpg) / (auto.mpg)) * 100)), "%"))
```

